---
title: "p8105_hw5_ps3395"
author: "Peng Su"
date: "2023-11-12"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(rvest)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 9, 
  fig.height = 6,
  out.width = "90%",
	fig.align = 'center'
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

# Problem 2

### Create a tidy dataframe

Writing functions to import all data and tidy the data.

```{r}
# get all data file names

file_names = 
  tibble(names = list.files("./data")) |>
  mutate(
    names = paste("./data", names, sep = "/")
  )

# read in the data for all subject

import = function(file_names) {

    read_csv(file_names) |>
    mutate(
      file_names = file_names
    )
}

# tidy the data

longitudinal_df = 
  bind_rows(map(file_names, import)) |>
  mutate(
    file_names = sub("^./data/", "", file_names),
    file_names = sub(".csv$", "", file_names)
  ) |>
  separate(file_names, into = c("arm","subject_ID"),sep = "_") |>
  pivot_longer(
    week_1:week_8,
    names_to = "time",
    values_to = "observation"
  ) |>
  select(subject_ID, everything()) |>
  janitor::clean_names()
  
```

### spaghetti plot

```{r}
# spaghetti plot

longitudinal_df |>
  group_by(subject_id) |>
  ggplot(aes(x = time, 
             y = observation, 
             group = as.factor(subject_id),
             color = as.factor(subject_id))) +
  geom_point(alpha = 0.3)+
  geom_line(alpha = 0.5)+
  facet_grid(. ~ arm) +
  labs(x = "Week", 
       y = "Observation",
       title = "Observations on each subject over time",
       caption = "Fig 1, The spaghetti plot shows the observations on each subject over time, where different color indicate different subjects",
       color = "Subject"
       ) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

The spaghetti plot displayed the observations on each subject over time, from Fig 1, it can be noticed that participants' observation data in the experimental arm and the control arm had small within-group differences, which the data fluctuated in a similar range. However, significant inter-group differences could be seen by comparing the experimental arm and the control arm. The observed data of the experimental group showed an overall increasing trend over time, while the data of the control group showed a relatively stable change.

# Problem 3

Generate 5000 datasets from the normal model with fixed simple size and sigma.

```{r}
# withing function for the mean and t.test when mu = 0
set.seed(12138)

n = 30
sigma = 5

sim_t.test = function(mu) {
  
  sim_data = tibble(
    x = rnorm(n, mean = mu, sd = sigma),
  )

  t.test(sim_data, mu = 0, conf.level = 0.05) |>
    broom::tidy() |>
    select(estimate, p.value)
}

#Generate 5000 datasets from the model for different mu

mu = c(0:6)

results =
  map_df(mu, ~ replicate(5000, sim_t.test(.x), simplify = FALSE)) |>
  mutate(
    mu = as.factor(rep(mu, each = 5000)),
    reject = case_when(
      p.value < 0.05 ~ "reject",
      p.value >= 0.05 ~ "fail_to_reject"
    )
    ) 

# plot showing the proportion of times the null was rejected

power_df = 
  results |>
  group_by(mu) |>
  count(reject) |>
  ungroup()|>
  pivot_wider(names_from = reject, values_from = n) |>
  mutate(
    fail_to_reject = ifelse(is.na(fail_to_reject), 0, fail_to_reject),
    power = reject / (fail_to_reject + reject)
  )

power_df |>
  ggplot(aes(x = mu, 
             y = power,
             group = 1)) +
  geom_point(aes(color = mu),alpha = 0.3)+
  geom_line(alpha = 0.5)+
  labs(x = "True mean", 
       y = "Power",
       title = "Proportion of times the null was rejected",
       caption = "Fig 2, The proportion of times the null was rejected under different samples with different true means",
       color = "True mean"
       )

#plot showing the average estimate of mu hat on the y axis and the true value of mu on the x axis.

average_muhat = 
  results |>
  group_by(mu) |>
  mutate(
    average_muhat = mean(estimate)
  ) 

average_muhat |>
  ggplot(aes(x = mu, 
             y = average_muhat,
             group = 1)) +
  geom_point(aes(color = mu),alpha = 0.3)+
  geom_line(alpha = 0.5)+
  labs(x = "True mean", 
       y = "average estimate of mu hat",
       title = "Comparison of average estimate of mu hat and the true mean",
       caption = "Fig 3, The comparison of average estimate of mu hat and the true means",
       color = "True mean"
       )

# making the second plot

average_muhat = 
  results |>
  group_by(mu) |>
  filter(
    reject == "reject"
  ) |>
  mutate(
    average_muhat = mean(estimate)
  ) 

average_muhat |>
  ggplot(aes(x = mu, 
             y = average_muhat,
             group = 1)) +
  geom_point(aes(color = mu),alpha = 0.3)+
  geom_line(alpha = 0.5)+
  labs(x = "True mean", 
       y = "average estimate of mu hat",
       title = "Comparison of average estimate of mu hat and the true mean",
       caption = "Fig 4, The comparison of average estimate of mu hat only in samples for which the null was rejected and the true means",
       color = "True mean"
       )
```

As in one-sample t-test, when the difference between the real $\mu$ and the $\mu$ ($\mu = 0$) of the null hypothesis is large, it represents a large effect size. From Fig 2, where the X-axis represents the true $\mu$ value, and the Y-axis represents the proportion of rejection of the null hypothesis (the power of the test). It is noticeable that when the difference between the true $\mu$ value and the test $\mu$ value (0) is large, which representing a large effect size, a higher proportion of rejection of the null hypothesis is appeared, in other words a higher statistical power, is usually observed.

By comparing Fig 3 and Fig 4, it can be seen that when the real $\mu$ have greater difference with the $\mu$ value (0) in the null hypothesis ($\mu$ = 4,5,6), the sample average of $\hat\mu$ across tests for which the null is rejected approximately equal to the true value of true $\mu$, but when the difference is small, difference between the average of $\hat\mu$ from the true value of $\mu$ is large. 

This may because for samples with a smaller gap between the true and assumed test $\mu$, more samples may be needed to observe significant differences, so the estimate of $\hat\mu$ may be more affected by extreme values, making it much different from the true $\mu$ value, at the same time for smaller differences, the test method may be less sensitive, making the estimate of $\hat\mu$ more likely to be affected by measurement error or randomness.